\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-15 trunk/arch.tex
Refined  some text. Ensured that some definitions (such as tau , lambda and pi) are at least mentioned early Reworked the synchronization with a focus on the current state flaw definition and proposed resolution

A prime motivation behind the \rx architecture was to design a
controller which could bring automated planning closer to low-level
control of the vehicle while maintaining its reactivity. Traditionally
planning has been deemed computationally expensive with planner
performance restricting robot reactivity
\cite{ghallab04,Dias:2003ua}. In control architectures where task
planning is embedded, planning consequently remains at an abstract
level with the assumption that it would otherwise impede system
reactiveness and that the environment can change at a faster rate than
the planner can plan for it. In such a situation, the agent may thrash
if the internal state of the plan gets out of synch with the actual
state of the world. As a result the planning problem managed {\em in
  situ} remains fairly detached from low-level control. Most
adaptations at the lower-level reactive layers in contrast are managed
by specialized components (for example GeNoM \comment{} and CLARATy
\comment{}) that rely on different representations for modeling local
control behavior or worse do not have any explicit formal agent
behavior model. The consequent different techniques for specifying
each layer in the architecture results in duplication of effort and a
diffusion of knowledge leading to significant design and integration
issues \cite{DS1report}.  \texttt{IDEA} mitigated most of these issues
by tightly integrating planning and execution in a single
representational and computational framework with a unified
declarative model. Our approach with \texttt{T-REX} builds on this
design methodology albeit in a substantially systematic manner

@@

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as
  each reactor declares its planning look-ahead and expected latency
  \comment{$\pi \&  \lambda$ both defined for the first time} before
  being able to produce plan to the agent depending on where it is
  located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

@@

This {\em synchronization} process propagates up through the reactor
dependency graph bottom up at every {\em tick} ensuring that all the
reactors have a consistent view of the world. Conversely each reactor
can request a future change on one of its {\em external} state
variables which will be transfromed as an {\em internal} goal to the
owner. Such \emph{goal dispatching} is expected to produce a
deliberation phase for the receiving reactor which can cyclically and
in turn, produce a plan impacting its {\em external} state variables
that will propagate down following the same mechanism.

@@

We first show how a reactor can track and identify state when it does
not have any compelling need to deliberate with no future goals. Even
in this condition, \rx requires each reactor to produce at every tick
its {\em internal} state. For those reactors not owning {\em internal}
timelines, \rx makes sure that {\em external} observations are taken
into account for every tick as shown in \cite{py10}. Such a
requirement is enforced to ensure that all reactors share a consistent
view of the world with advancing time.

@@

We first show how a reactor can track and identify its state in the
context where it does not have any compelling need to deliberate. We
assume that we have a reactor that has no future goal. \comment{Need
  to develop the reason why it is a necessity: primarily to provide
  its internal state to whoever observe it but also simply to ensure
  that its current representation of the world is up to date and still
  consistent}

We introduce this requirement as a new type of flaw for the \eu
framework. This flaw forces the reactor to identify its internal state
for the current tick. By using this flaw one can describe the
synchronization process as:

@@

An internal state flaw can be resolved using one of the following
choices which are evaluated in order:

\item Extend the previous state value to end after this tick (\ie
  restrict its end time to $[\tau+1, \infty)$). \comment{first use of
    $\tau$, so needs to be defined earlier}

All of these choices are evaluated sequentially until a consistent
solution with no flaws at this tick, is identified. In order to do so
we need to make the assumption that the current state value of an
internal timeline does not depend on the future or more accurately
that any choices made during this synchronization will not lead to a
future inconsistency. Such an assumption impacts the set of possible
domains the reactor can support while remaining complete. Take for
example this rule from our \texttt{Shopping} model :

>>>>

A prime motivation behind the \rx architecture was to design a
controller which could bring automated planning closer to low-level
control of the vehicle while maintaining its reactivity. Traditionally
planning has been deemed computationally expensive with planner
performance restricting robot reactivity
\cite{ghallab04,Dias:2003ua}. In control architectures where task
planning is embedded, planning consequently remains at an abstract
level with the assumption that it would otherwise impede system
reactiveness and that the environment can change at a faster rate than
the planner can plan for it. In such a situation, the agent may thrash
if the internal state of the plan gets out of synch with the actual
state of the world. As a result the planning problem managed {\em in
  situ} remains fairly detached from low-level control. Most
adaptations at the lower-level reactive layers in contrast are managed
by specialized components (for example LAAS architecture
\cite{alami:1998p820,Ingrand07} and CLARATy \cite{Nesnas:2003do}
) that rely on different representations for modeling local
control behavior or worse do not have any explicit formal agent
behavior model. The consequent different techniques for specifying
each layer in the architecture results in duplication of effort and a
diffusion of knowledge leading to significant design and integration
issues \cite{DS1report}.  \texttt{IDEA} mitigated most of these issues
by tightly integrating planning and execution in a single
representational and computational framework with a unified
declarative model. Our approach with \texttt{T-REX} builds on this
design methodology albeit in a substantially systematic manner

@@

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as
  each reactor declares its planning look-ahead $\pi$  and expected
  latency $\lambda$ 
  % \comment{$\pi \&  \lambda$ both defined for the first time} 
  before
  being able to produce plan to the agent depending on where it is
  located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

@@

This {\em synchronization} process propagates the current state
information at the execution frontier $\tau$ throughout all the 
reactors. The sequencing of the reactors synchronization is identified
by traversing the reactor graph from the least dependents reactors --
the ones with no {\em external} timelines -- to the most dependent
ones which typicallu do not provide {\em internal} timelines.
Conversely each reactor can request a future change on one of 
its {\em external} state variables which will be transfromed as 
an {\em internal} goal to the owner. 
Such \emph{goal dispatching} is expected to produce a
deliberation phase for the receiving reactor which can cyclically and
in turn, produce a plan impacting its {\em external} state variables' 
future that will propagate down following the same mechanism.

@@

We first show how a reactor can track and identify its state in the
context where it does not have any compelling need to deliberate. We
assume that we have a reactor that has no future goal.Even
in this condition, \rx requires each reactor to produce at every tick
its {\em internal} state. For those reactors not owning {\em internal}
timelines, \rx makes sure that {\em external} observations are taken
into account for every tick as shown in \cite{py10}. Such a
requirement is enforced to ensure that all reactors share a consistent
view of the world with advancing time.

@@

We introduce this requirement as a new type of flaw for the \eu
framework defined as follow: 
\begin{definition}
  \label{def:csf}
  A {\em current state flaw} indicate that one of the reactor {\em
    internal} timelines does not provide a fully defined state value
  for the execution frontier $\tau$. Such flaw can be due to one 
  of the 2 following reasons:
  \begin{itemize}
  \item The absence of a token in this timeline which overlaps the
    instant $\tau$. 
  \item An open decision between terminating the last past token for
    this timeline or pushing its end in the future.
  \end{itemize}
\end{definition}

The core idea bewteen this flaw is to ensure that at any single tick
the {\em synchronization} solver will identifies all the {\em
  internal} timelines for which the current state value is yet to be
identified. We handle their resolution based on following
proposition. 
\begin{proposition}
  \label{prop:csf:resolve}
  A {\em current state flaw} can  be resolved using one of the
  following choices which are eveluated accroding to this order:
\begin{enumerate}

\item Extend the previous state value to end after the execution 
  frontier  (\ie restrict its end time to $[\tau+1, \infty)$). %\comment{first use of
%    $\tau$, so needs to be defined earlier}

\end{enumerate}
\end{proposition}
Intuitively the first choice enforces an intertail value asumption
where the reactor will tend to maintain its current as long as the
model and external observations allow it. The second choice 
favorise the adavnce of the reactor current plan if any; this direct
the {\em synchronization} decision toward assuming that its current
plan is properly executed. Finally if none of these two choices did
provide a valid solution -- meaning that both lead to an inconsittant
plan -- the reactor tries all the possible state values remaining for
this timeline until it finds one that allows it to identify its
state as required by \rx architecture.

This flaw forces the reactor to identify its internal state
for the current tick. By using this flaw one can describe the
synchronization process as:

@@

\item Resolve the current state value of each internal timelines based
  on Proposition \ref{prop:csf:resolve}.
\end{enumerate}

The steps 2 \& 3 are handled by the \eu solver from algorithm
\ref{alg:europa:solve}. We only enforce that the {\em current state
  flaws} are selected last by the algorithm in order to follow the
sequencing given above. As {\em synchronization} focusses on the
execution frontier wee also filter out all flaws that are outside of
the current tick scope. This mean that if this flaw relates to a token
that is {\em necessarily} ending before $\tau$ or starting after 
$\tau+1$ then the solver will ignore it. This helps to reduce the
scope of the synchronization to exclusively the current state of the
reactor and reduce dramatically the search space for the {\em
  synchronization} solver. In order to do so we need to make the
assumption that the current state value of an
{\em internal} timeline does not depend on the future or more accurately
that any choices made during this synchronization will not lead to a
future inconsistency. Such an assumption impacts the set of possible 
domains the reactor can support while remaining complete. Take for 
example this rule from our \texttt{Shopping} model :
